Matching Bullets

### 2015-12-12 imaging-paper.Rnw
little edits

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Rectangles in the back identify a line on one of the bullets.  Matching striation marks are indicated by color filled rectangles and arked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

\begin{figure}[hbtp]
%\centering
%\begin{subfigure}[t]{\linewidth}
%\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(smoothfactor~., labeller="label_both") +
    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
              colour="grey50") +
    geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
@
%'     }
%' \end{subfigure}    
%' \begin{subfigure}[t]{\linewidth}
%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
%' <<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
%' get_peaks(br111.loess$data, smoothfactor = 5)$plot
%' @
%'     }
%' \end{subfigure}
%' \begin{subfigure}[t]{\linewidth}
%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
%' <<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
%' get_peaks(br111.loess$data, smoothfactor = 65)$plot
%' @
%'     }
%' \end{subfigure}
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 15, and 35, respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 35 might over-smoothe and cause the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
\end{figure}

@@

bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
qplot(data=bm, value, reorder(bullet, value/n), size=I(2.5), 
      shape=prediction, colour=prediction, alpha=I(0.6)) + 
  theme_bw()  + ylab("") + 
#  scale_shape_manual(values=c(1,2)) +
  scale_x_continuous("Number of correctly predicted land-to-land matches", 
                     breaks = 3*0:4, limits=c(0,12),
                     minor_breaks = 1:11) +
  theme(legend.position = "bottom")
@
\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree. Both results for the tree and the random forest are shown. For each bullet at least four lands are correctly matched. Unknown bullets B and M are the hardest for the algorithm to match.}
\end{figure}
%
However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger test case. 
Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on 0.5 is not necessary. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis). Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only 12 cases result in a different decision.

\begin{figure}[hbtp]
<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
qplot(pred, forest, geom="jitter", data=bstats, colour=match) + theme_bw() +
  theme(legend.position="bottom") +
  geom_vline(xintercept=0.5, colour="grey50") +
  geom_hline(yintercept=0.5, colour="grey50") + 
  xlab("Predictions from decision tree") +
  ylab("Predictions from random forest")   +
  scale_colour_brewer(palette="Paired")
@
\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical). Each point corresponds to a land-to-land comparison. Known matches are indicated in blue. Matches predicted by the decision tree are on the right half of the plot, matches predicted by the random forest are in the upper half. }
\end{figure}

>>>>

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Rectangles in the back identify a line on one of the bullets.  Matching striation marks are indicated by color filled rectangles and arked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

\begin{figure}[hbtp]
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(smoothfactor~., labeller="label_both") +
    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
              colour="grey50") +
    geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
@
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 15, and 35, respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 35 might over-smoothe and cause the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
\end{figure}

@@

bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
qplot(data=bm, value, reorder(bullet, value/n), size=I(2.5), 
      shape=prediction, colour=prediction, alpha=I(0.6)) + 
  theme_bw()  + ylab("") + 
#  scale_shape_manual(values=c(1,2)) +
  scale_x_continuous("Number of correctly predicted land-to-land matches", 
                     breaks = 3*0:4, limits=c(0,12),
                     minor_breaks = 1:11) +
  theme(legend.position = "bottom")
@
\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12). Both results for the tree and the random forest are shown. For each bullet at least four lands are correctly matched. Unknown bullets B and M are the hardest for the algorithm to match.}
\end{figure}
%
However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger test case. 
Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on 0.5 is not necessary. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis). Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only 12 cases result in a different decision.

\begin{figure}[hbtp]
<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
      shape=match) + theme_bw() +
  theme(legend.position="bottom") +
  geom_vline(xintercept=0.5, colour="grey50") +
  geom_hline(yintercept=0.5, colour="grey50") + 
  xlab("Predictions from decision tree") +
  ylab("Predictions from random forest")   +
  scale_colour_brewer(palette="Paired")
@
\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical). Each point corresponds to a land-to-land comparison. Known matches are indicated in blue. Matches predicted by the decision tree are on the right half of the plot, matches predicted by the random forest are in the upper half. }
\end{figure}
